---
title: "Data Preparation and App Deployment"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Preparation and App Deployment}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(gsm.app)
library(dplyr)
library(purrr)
```

# Introduction

The {gsm.app} package creates interactive dashboards for exploring Good Statistical Monitoring (GSM) Key Risk Indicator (KRI) assessments in clinical trials. This vignette provides a comprehensive overview on how to prepare data for use with the {gsm.app} Shiny application and how to deploy an app.

The {gsm} suite of packages leverages Key Risk Indicators (KRIs) and thresholds to conduct study-level, country-level and site-level Risk Based Monitoring for clinical trials. {gsm.app} package tailored to the {gsm} analysis data model that connects risk signals with the underlying data.

Preparing data for use with {gsm.app} involves a standardized, reproducible pipeline leveraging the {gsm.core}, {gsm.mapping}, {gsm.kri}, and {gsm.reporting} packages. The process consists of the following high-level steps:

1.  [Set Up Data Sources](https://gilead-biostats.github.io/gsm.core/articles/DataModel.html#raw-and-mapped-data).

    Define and load your raw clinical and operational data.

2.  [Create Mapping Workflows](https://gilead-biostats.github.io/gsm.core/articles/Cookbook.html#example-3---study-level-reporting-workflows)

    Specify and apply mapping workflows to transform raw data into standardized mapped data.

3.  [Generate Mapped Data Layer](https://gilead-biostats.github.io/gsm.core/articles/Cookbook.html#automate-data-ingestion-using-ingest-and-combinespecs)

    Use mapping specifications to ingest, filter, and join data, producing a harmonized mapped data layer.

4.  [Run Analysis Workflows](https://gilead-biostats.github.io/gsm.core/articles/DataAnalysis.html)

    Apply analysis workflows to calculate Key Risk Indicators (KRIs) and perform statistical modeling.

5.  [Create Reporting Data](https://gilead-biostats.github.io/gsm.reporting/articles/DataReporting.html#step-by-step-reporting-workflow)

    Transform analysis results into reporting-ready data frames for use in dashboards and reports. The reporting data model is documented in detail at [Reporting Data](https://gilead-biostats.github.io/gsm.core/articles/DataModel.html#reporting-data).

At this point, you should have all required data structures to deploy the app.

The {gsm.app} requires five main data frames to operate:

1.  **`dfAnalyticsInput`** - Participant-level metric data (stacked across metrics)
2.  **`dfBounds`** - Statistical bounds for flagging. Set of predicted percentages/rates and upper- and lower-bounds across the full range of sample sizes/total exposure values for reporting.
3.  **`dfGroups`** - Group-level metadata (sites, countries, study)
4.  **`dfMetrics`** - Metric-specific metadata for use in charts and reporting.
5.  **`dfResults`** - KRI assessment results. A stacked summary of analysis pipeline output.

# Data Fetching Function (`fnFetchData`)

The gsm app allows you to provide custom data fetching functions that read and subset data from your local file system or database. The [fnFetchData](https://gilead-biostats.github.io/gsm.app/reference/sample_fnFetchData.html) function is essential for the {gsm.app} to provide drill-down functionality for domain-specific data. This function tells the app how to fetch domain data and should return filtered data based on the parameters passed from the app interface.

The function definition can be customized to suite your system and data needs, however there are several factors that should be accounted for when creating a `fnFetchData` function

The `fnFetchData` function should accept the following parameters and return a filtered data frame:

-   `strDomain`: Character string specifying the data domain (e.g., "AE", "LB", "DM")
-   `strGroupLevel`: Character string specifying the group level (e.g., "Site", "Country")
-   `strGroupID`: Character string specifying the specific group ID to filter on
-   `strSubjectID`: Character string specifying the subject ID (optional, may be NULL)

### Basic Implementation

Here's a basic template for a data fetching function assuming your data stored locally:

```{r create-fnfetchdata, eval = FALSE}
fnFetchData <- function(strDomain,
                      strGroupID = NULL,
                      strGroupLevel = "Site",
                      strSubjectID = NULL,
                      dSnapshotDate = NULL) {
  
  # Define base path to your data files
  data_path <- "path/to/your/data"
  
  # Construct file path based on domain
  file_path <- file.path(data_path, paste0(strDomain, ".csv"))
  
  # Read the data
  dfDomain <- read.csv(file_path, stringsAsFactors = FALSE)

  #rename columns, if needed
  dfDomain <- dfDomain %>%
    dplyr::rename(
      "SubjectID" = "USUBJID"
    )

  # Apply group-level filtering, if specified
  if (!is.null(strGroupID) && "GroupID" %in% colnames(dfDomain)) {
    dfDomain <- dplyr::filter(dfDomain, .data$GroupID == strGroupID)
  }

  # Apply subject-level filtering, if specified
  if (!is.null(strSubjectID) && "SubjectID" %in% colnames(dfDomain)) {
    dfDomain <- dplyr::filter(dfDomain, .data$SubjectID == strSubjectID)
  }

  return(dfDomain)
}
```

### Advanced Implementation with Multiple Data Sources

For more complex scenarios where you have multiple data sources or need custom logic:

```{r eval=FALSE}

advanced_fetch_data <- function(strDomain, 
                                strGroupLevel, 
                                strGroupID, 
                                strSubjectID = NULL) {
  # Define data source mappings
  data_sources <- list(
    "AE" = "adverse_events.csv",
    "LB" = "lab_data.csv", 
    "DM" = "demographics.csv",
    "VS" = "vital_signs.csv"
  )
  
  # Base data directory
  base_path <-  "path/to/your/data"
  
  # Get the appropriate file for the domain
  if (!strDomain %in% names(data_sources)) {
    warning(paste("Unknown domain:", strDomain))
    return(data.frame())
  }
  
  file_name <- data_sources[[strDomain]]
  full_path <- file.path(base_path, file_name)
  
  # Read data with error handling
  tryCatch({
    data <- read.csv(full_path, stringsAsFactors = FALSE)
  }, error = function(e) {
    warning(paste("Error reading file:", full_path, "-", e$message))
    return(data.frame())
  })
  
  # Domain-specific preprocessing
  if (strDomain == "AE") {
    # Ensure required columns exist for adverse events
    required_cols <- c("SubjectID", "AETERM", "AESEV", "AEREL")
    missing_cols <- setdiff(required_cols, names(data))
    if (length(missing_cols) > 0) {
      warning(paste("Missing required columns for AE domain:", paste(missing_cols, collapse = ", ")))
    }
  }
  
  # Apply group-level filtering, if specified
  if (!is.null(strGroupID) && "GroupID" %in% colnames(dfDomain)) {
    dfDomain <- dplyr::filter(dfDomain, .data$GroupID == strGroupID)
  }

  # Apply subject-level filtering, if specified
  if (!is.null(strSubjectID) && "SubjectID" %in% colnames(dfDomain)) {
    dfDomain <- dplyr::filter(dfDomain, .data$SubjectID == strSubjectID)
  }
  
  return(data)
}
```

## Data Counting Function (`fnCountData`)

The Domain Summary tab in gsm.app provides information on number of records for each domain. When working with large datasets, retrieving only the count of records is often significantly faster than fetching the full dataset. However, if your `fnFetchData()` function is already fast, `fnCountData` allows you to construct a simple data counter from your data fetcher as it can be just a wrapper around `fnFetchData` . The [`fnCountData`](https://gilead-biostats.github.io/gsm.app/reference/ConstructDataCounter.html) function should accept the `fnFetchData` function as a parameter, but return an integer count instead of a data frame.

### Implementation

```{r eval=FALSE}
fnCountData <- function(fnFetchData) {
  # Reuse the fetch function and count rows
  data <- fnFetchData(strDomain, strGroupLevel, strGroupID, strSubjectID)
  return(nrow(data))
}
```

# App Deployment

Once you have prepared all the required data structures and functions, you can pass them to `run_gsm_app()`.

For development and testing, run the app locally: run_gsm_app() creates a Shiny app to explore a set of clinical trial data. The app facilitates exploration of the data by allowing the user to click to dive deeper into aspects of the data.

```{r local-deployment, eval = FALSE}
# Run the app with custom data function
run_gsm_app(
  dfAnalyticsInput = dfAnalyticsInput,
  dfBounds = dfBounds,
  dfGroups = dfGroups,
  dfMetrics = dfMetrics,
  dfResults = dfResults,
  fnFetchData = fnFetchData,
  fnCountData = fnCountData(fnFetchData),
  chrDomains = c(AE = "Adverse Events", DATACHG = "Data Changes", DATAENT = "Data Entry",
    ENROLL = "Enrollment", LB = "Lab", PD = "Protocol Deviations", QUERY = "Queries",
    STUDCOMP = "Study Completion", SUBJ = "Subject Metadata", SDRGCOMP =
    "Treatment Completion"),
  lPlugins = NULL,
  strTitle = ExtractAppTitle(dfGroups),
  strFavicon = "angles-up",
  strFaviconColor = "#FF5859"
)
```

# 
