<!-- filepath: c:\Users\schildress\dev\gsm.app\vignettes\data-fetching-functions.Rmd -->
---
title: "Writing Custom Data Fetching Functions for GSM App"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Writing Custom Data Fetching Functions for GSM App}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gsm.app)
```

## Overview

The GSM app allows you to provide custom data fetching functions that read and subset data from your local file system or database. This vignette demonstrates how to write functions that can be passed to `run_gsm_app()` via the `fnFetchData` and `fnCountData` parameters.

## Data Fetching Function (`fnFetchData`)

The `fnFetchData` function should accept the following parameters and return a filtered data frame:

- `strDomain`: Character string specifying the data domain (e.g., "AE", "LB", "DM")
- `strGroupLevel`: Character string specifying the group level (e.g., "Site", "Country") 
- `strGroupID`: Character string specifying the specific group ID to filter on
- `strSubjectID`: Character string specifying the subject ID (optional, may be NULL)

### Basic Implementation

Here's a basic template for a data fetching function:

```{r eval=FALSE}
my_fetch_data <- function(strDomain, strGroupLevel, strGroupID, strSubjectID = NULL) {
  # Define base path to your data files
  data_path <- "path/to/your/data"
  
  # Construct file path based on domain
  file_path <- file.path(data_path, paste0(strDomain, ".csv"))
  
  # Check if file exists
  if (!file.exists(file_path)) {
    warning(paste("Data file not found:", file_path))
    return(data.frame())
  }
  
  # Read the data
  data <- read.csv(file_path, stringsAsFactors = FALSE)
  
  # Apply group-level filtering
  if (!is.null(strGroupLevel) && strGroupLevel != "" && strGroupLevel %in% names(data)) {
    if (!is.null(strGroupID) && strGroupID != "") {
      data <- data[data[[strGroupLevel]] == strGroupID, ]
    }
  }
  
  # Apply subject-level filtering if specified
  if (!is.null(strSubjectID) && strSubjectID != "" && "SubjectID" %in% names(data)) {
    data <- data[data$SubjectID == strSubjectID, ]
  }
  
  return(data)
}
```

### Advanced Implementation with Multiple Data Sources

For more complex scenarios where you have multiple data sources or need custom logic:

```{r eval=FALSE}
advanced_fetch_data <- function(strDomain, strGroupLevel, strGroupID, strSubjectID = NULL) {
  # Define data source mappings
  data_sources <- list(
    "AE" = "adverse_events.csv",
    "LB" = "lab_data.csv", 
    "DM" = "demographics.csv",
    "VS" = "vital_signs.csv"
  )
  
  # Base data directory
  base_path <- "C:/clinical_data"
  
  # Get the appropriate file for the domain
  if (!strDomain %in% names(data_sources)) {
    warning(paste("Unknown domain:", strDomain))
    return(data.frame())
  }
  
  file_name <- data_sources[[strDomain]]
  full_path <- file.path(base_path, file_name)
  
  # Read data with error handling
  tryCatch({
    data <- read.csv(full_path, stringsAsFactors = FALSE)
  }, error = function(e) {
    warning(paste("Error reading file:", full_path, "-", e$message))
    return(data.frame())
  })
  
  # Domain-specific preprocessing
  if (strDomain == "AE") {
    # Ensure required columns exist for adverse events
    required_cols <- c("SubjectID", "AETERM", "AESEV", "AEREL")
    missing_cols <- setdiff(required_cols, names(data))
    if (length(missing_cols) > 0) {
      warning(paste("Missing required columns for AE domain:", paste(missing_cols, collapse = ", ")))
    }
  }
  
  # Apply group filtering with validation
  if (!is.null(strGroupLevel) && strGroupLevel != "") {
    if (strGroupLevel %in% names(data)) {
      if (!is.null(strGroupID) && strGroupID != "") {
        # Handle both exact matches and pattern matching
        if (strGroupLevel == "Site") {
          data <- data[data[[strGroupLevel]] == strGroupID, ]
        } else {
          # For other group levels, you might want pattern matching
          data <- data[grepl(strGroupID, data[[strGroupLevel]], ignore.case = TRUE), ]
        }
      }
    } else {
      warning(paste("Group level column not found:", strGroupLevel))
    }
  }
  
  # Subject filtering
  if (!is.null(strSubjectID) && strSubjectID != "") {
    if ("SubjectID" %in% names(data)) {
      data <- data[data$SubjectID == strSubjectID, ]
    } else {
      warning("SubjectID column not found in data")
    }
  }
  
  # Remove any rows with all NA values
  data <- data[rowSums(is.na(data)) != ncol(data), ]
  
  return(data)
}
```

## Data Counting Function (`fnCountData`)

The `fnCountData` function should accept the same parameters as `fnFetchData` but return an integer count instead of a data frame.

### Basic Implementation

```{r eval=FALSE}
my_count_data <- function(strDomain, strGroupLevel, strGroupID, strSubjectID = NULL) {
  # Reuse the fetch function and count rows
  data <- my_fetch_data(strDomain, strGroupLevel, strGroupID, strSubjectID)
  return(nrow(data))
}
```

### Optimized Implementation

For better performance, especially with large datasets, you might want to implement counting without loading all data:

```{r eval=FALSE}
optimized_count_data <- function(strDomain, strGroupLevel, strGroupID, strSubjectID = NULL) {
  # Define data source mappings (same as fetch function)
  data_sources <- list(
    "AE" = "adverse_events.csv",
    "LB" = "lab_data.csv", 
    "DM" = "demographics.csv",
    "VS" = "vital_signs.csv"
  )
  
  base_path <- "C:/clinical_data"
  
  if (!strDomain %in% names(data_sources)) {
    return(0L)
  }
  
  file_name <- data_sources[[strDomain]]
  full_path <- file.path(base_path, file_name)
  
  if (!file.exists(full_path)) {
    return(0L)
  }
  
  # For very large files, you might want to use data.table or other optimized reading
  tryCatch({
    # Read only necessary columns for filtering
    if (requireNamespace("data.table", quietly = TRUE)) {
      # Use data.table for efficient reading and filtering
      data <- data.table::fread(full_path)
      
      # Apply filters
      if (!is.null(strGroupLevel) && strGroupLevel != "" && strGroupLevel %in% names(data)) {
        if (!is.null(strGroupID) && strGroupID != "") {
          data <- data[data[[strGroupLevel]] == strGroupID]
        }
      }
      
      if (!is.null(strSubjectID) && strSubjectID != "" && "SubjectID" %in% names(data)) {
        data <- data[data[["SubjectID"]] == strSubjectID]
      }
      
      return(nrow(data))
    } else {
      # Fallback to base R
      data <- read.csv(full_path, stringsAsFactors = FALSE)
      
      # Apply same filtering logic as fetch function
      if (!is.null(strGroupLevel) && strGroupLevel != "" && strGroupLevel %in% names(data)) {
        if (!is.null(strGroupID) && strGroupID != "") {
          data <- data[data[[strGroupLevel]] == strGroupID, ]
        }
      }
      
      if (!is.null(strSubjectID) && strSubjectID != "" && "SubjectID" %in% names(data)) {
        data <- data[data$SubjectID == strSubjectID, ]
      }
      
      # Remove rows with all NA
      data <- data[rowSums(is.na(data)) != ncol(data), ]
      
      return(nrow(data))
    }
  }, error = function(e) {
    warning(paste("Error counting data:", e$message))
    return(0L)
  })
}
```

## Using the Functions with `run_gsm_app()`

Once you've defined your data fetching functions, pass them to the GSM app:

```{r eval=FALSE}
# Run the app with custom data functions
run_gsm_app(
  fnFetchData = my_fetch_data,
  fnCountData = my_count_data
)

# Or with the advanced versions
run_gsm_app(
  fnFetchData = advanced_fetch_data,
  fnCountData = optimized_count_data
)
```

## Best Practices

1. **Error Handling**: Always include proper error handling to prevent the app from crashing
2. **Data Validation**: Validate that required columns exist in your data
3. **Performance**: For large datasets, consider using `data.table` or similar packages for efficient data operations
4. **Consistency**: Ensure both `fnFetchData` and `fnCountData` apply the same filtering logic
5. **Documentation**: Document the expected data structure and column names
6. **Testing**: Test your functions with various parameter combinations before deployment

## Example Data Structure

Your data files should typically include these columns:

- `SubjectID`: Unique subject identifier
- Group level columns (e.g., `Site`, `Country`, `Study`)
- Domain-specific data columns
- Any additional metadata columns

Example structure for adverse events data:
```
SubjectID | Site | Country | AETERM | AESEV | AEREL | AESTDT
001       | 101  | USA     | Nausea | Mild  | Possible | 2023-01-15
002       | 101  | USA     | Headache | Moderate | Unlikely | 2023-01-16
```

This structure allows the filtering functions to work properly with the