---
title: "Plugins"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plugins}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gsm.app)
```

# Introduction

The plugin system provides a flexible way to tailor applications to specific needs by incorporating additional features and dependencies.

# Defining a Plugin

A plugin is defined by a folder containing:

- A YAML file with metadata and dependency information.
- Optional R scripts that define the necessary UI and server functions for the plugin.

For example, this is the`AE.yml` file for our sample Adverse Events plugin:

```yaml
meta:
  Type: Plugin
  ID: AE
  Name: Adverse Events
shiny:
  UI: mod_AE_UI
  Server: mod_AE_Server
domains:
  - AE
  - SUBJ
packages:
  - name: ggplot2
  - name: safetyCharts
    remote: url::https://safetygraphics.r-universe.dev/src/contrib/safetyCharts_0.4.0.tar.gz
```

## meta

The `meta` section must contain `Type: Plugin`, and `ID` that matches the name of the file, and a `Name:`. The `Name:` appears in the resulting app as the label for the tab containing the plugin.

## shiny

The `shiny` section specifies the names of the plugin module's UI and server functions. These functions can come from a package (included in `packages`, see below), or be defined in one or more R scripts in the plugin folder.

## domains

The `domains` section specifies the data domains that are used by the plugin. These domains will be loaded by the app, in addition to any domains specified in the `chrDomains` argument to `run_gsm_app()`. The domains can be any of the following:
- `AE` (Adverse Events)
- `DATACHG` (Data Changes)
- `DATAENT` (Data Entry)
- `ENROLL` (Enrollment)
- `LB` (Lab)
- `PD` (Protocol Deviations)
- `QUERY` (Queries)
- `STUDCOMP` (Study Completion)
- `SUBJ` (Subject Metadata)
- `SDRGCOMP` (Treatment Completion)

## required_inputs

The `required_inputs` section lists any inputs ("Metric", "Site", "Participant", or "Domain") that must have a non-empty value before the plugin can load. "None" and "All" count as "empty" for this check. If the user has not set a value for that input, the app will display a placeholder instructing the user to make a selection.

## packages

The `packages` section specifies packages used by the plugin. This field is particulaly important for deploying the app or using the app with a CI/CD system.

# Updating Inputs Inside a Plugin

If plugins have any of these arguments, the corresponding reactive will be passed to the plugin:

- `rctv_strMetricID` (a `shiny::reactiveVal()` connected to the "Metric" input)
- `rctv_strSiteID` (a `shiny::reactiveVal()` connected to the "Site" input)
- `rctv_strSubjectID` (a `shiny::reactiveVal()` connected to the "Participant" input)
- `rctv_strDomainID` (a `shiny::reactiveVal()` connected to the "Domain" input)
- `rctv_dfDOMAIN` (eg rctv_dfAE, one for each `domain` as defined above)

For the input-connected `reactiveVal()`s, setting a value will pass the value back up to the main app. For example, to set the `Participant` drop-down menu to subject "0008", call `rctv_strSubjectID("0008")` inside your plugin.

We recommend avoiding setting values to `NULL` from inside your app, to avoid issues with synchronization.

# Launching an App with a Plugin

Use `plugin_Read(strPath)` to read a plugin definition and prepare it for use in an app. Supply a list of one or more plugin definitions to the `run_gsm_app()` function through the `lPlugins` argument.

For example, this code launches an app with our sample AE plugin:

```{r launch, eval = FALSE}
aePlugin <- plugin_Read(system.file("plugins", "AE", package = "gsm.app"))
run_gsm_app(
  dfAnalyticsInput = gsm.app::sample_dfAnalyticsInput,
  dfBounds = gsm.app::sample_dfBounds,
  dfGroups = gsm.app::sample_dfGroups,
  dfMetrics = gsm.app::sample_dfMetrics,
  dfResults = gsm.app::sample_dfResults,
  fnFetchData = sample_fnFetchData,
  lPlugins = list(aePlugin)
)
```

Note that `aePlugin` must be supplied inside an enclosing `list()`, to allow for multiple plugin definitions.

# Deploying an App with a Plugin

If you wish to deploy an app that uses one or more plugins using {renv} and/or {rsconnect}, you need to let the system know about the plugins used by your app. The `plugin_LoadDependencies()` function serves to alert such systems of your requirements.

```{r load-dependencies, eval = FALSE}
plugin_LoadDependencies(aePlugin)
```

# Plugins in CI/CD

To integrate plugins into a CI/CD pipeline, use the `plugin_InstallDependencySources()` function to install plugin dependencies. This function requires the {pak} package.

```{r install-sources, eval = FALSE}
plugin_InstallDependencySources(aePlugin)
```

See the [`shinyapps-deploy.yaml` file used by this package](https://github.com/Gilead-BioStats/gsm.app/blob/dev/.github/workflows/shinyapps-deploy.yaml) for an example use of this function.

If you wish to use something other than `pak::pak()` to install the dependencies, you can get a character vector of dependency sources with `plugin_GetDependencySources()`.
