---
title: "BR_NUMBER: BR_TITLE"
format: 
  html:
    lightbox: true
fig-cap-location: top
---

```{r}
#| echo: false
# This is crazy but seems to maybe work, inspired by https://github.com/r-lib/nanoparquet/blob/e881a75cf4930156625c4b728a17f3cc6cd86796/vignettes/articles/benchmarks.qmd#L22
my_path <- Sys.getenv("QUARTO_DOCUMENT_PATH")
if (my_path == "") {
  my_path <- getwd()
}
my_path <- fs::path(my_path)

if (fs::dir_exists(fs::path(my_path, "requirements"))) {
  br_path <- fs::path(my_path, "requirements")
} else if (fs::file_exists("requirement_issues.yml")) {
  br_path <- my_path
} else if (
  fs::file_exists("articles/requirements/requirement_issues.yml")
) {
  br_path <- fs::path("articles/requirements")
} else {
  br_path <- fs::path("vignettes/articles/requirements")
}
```

```{r}
#| echo: false
#| message: false
library(magrittr)
br_issue_number <- BR_NUMBER
brs <- yaml::read_yaml(fs::path(br_path, "requirement_issues.yml"))
br <- purrr::keep(
  brs,
  ~ .x$issue == br_issue_number
)[[1]]

img_path <- fs::path(br_path, "img")
all_imgs <- fs::dir_ls(img_path, recurse = TRUE, type = "file", glob = "*.png")
all_imgs <- fs::path_rel(all_imgs, start = my_path)
names(all_imgs) <- fs::path_ext_remove(fs::path_file(all_imgs))
```

## Description

[GitHub issue](https://github.com/Gilead-BioStats/gsm.app/issues/`r br$issue`)

`r br$description`

---

## Features & Bug Fixes

Explore the screenshot evidence presented here to determine if you agree that the app does what is described. 
Click images for larger versions. 
If you agree that the app does what is described, go to the linked GitHub issue for that feature or bug fix, and comment with your approval.
If you do *not* agree, go to the linked GitHub issue for that feature or bug fix, and comment with what you believe is incorrect.

```{r}
#| echo: false
#| output: asis
qmd_output <- purrr::map_chr(br$features, function(feature) {
  # Generate unique IDs for this feature's collapsible element using the issue number.
  feature_target_id <- paste0("folder-", feature$issue)
  feature_target_class <- paste0(".", feature_target_id, "-contents")
  
  if ("superseded" %in% feature$labels) {
    acceptance_criteria_qmd <- "This feature has been superseded by one or more later features. See the GitHub issue for details."
  } else {
    acceptance_criteria_qmd <- paste(
      purrr::imap_chr(feature$acceptance_criteria, function(ac, ac_idx) {
        # Generate unique IDs for the acceptance criteria collapsible element.
        ac_target_id <- glue::glue("folder-ac-{feature$issue}-{ac_idx}")
        ac_target_class <- paste0(".", ac_target_id, "-contents")
        
        ac_pad <- stringr::str_pad(ac_idx, width = 2, pad = "0")
        img_pattern <- glue::glue("^{feature$issue}-{ac_pad}-")
        relevant_img_names <- stringr::str_subset(names(all_imgs), img_pattern)
        
        # Generate Quarto markdown for images.
        img_qmd <- ""
        if (length(relevant_img_names) > 0) {
          img_qmd <- paste(
            purrr::map_chr(relevant_img_names, function(img_name) {
              caption <- stringr::str_remove(img_name, "^(\\d+-)+") %>%
                stringr::str_replace_all(c("_" = " ", "-" = ": "))
              img_path <- all_imgs[[img_name]]
              
              glue::glue(
                '![{caption}]({img_path}){{.lightbox}}'
              )
            }),
            collapse = "\n\n"
          )
        }
        
        # Generate Quarto markdown for the acceptance criteria folder.
        glue::glue(
          '
:::::: {{.folder .acceptance}}
::::: {{.folder-header data-bs-toggle="collapse" data-bs-target="{ac_target_class}" aria-expanded="false" aria-controls="{ac_target_id}"}}
:::: {{.folder-title-container}}
{ac}
::::
:::: {{.folder-btn-toggle}}
::: {{.folder-toggle}}
:::
::::
:::::
::::: {{#{ac_target_id} .{sub("^[.]", "", ac_target_class)} .folder-collapse .collapse}}
:::: {{.folder-body-container}}
{img_qmd}
::::
:::::
::::::
'
        )
      }),
      collapse = "\n\n"
    )
  }

  # Generate Quarto markdown for the main feature folder.
  glue::glue(
    '
::::::::: {{#anchor-{feature$issue} .folder .feature}}
:::::::: {{.folder-header .collapsed data-bs-toggle="collapse" data-bs-target="{feature_target_class}" aria-expanded="false" aria-controls="{feature_target_id}"}}
::::::: {{.folder-title-container}}
{feature$issue}: {feature$title}
:::::::
::::::: {{.folder-btn-toggle}}
:::::: {{.folder-toggle}}
::::::
:::::::
::::::::
:::::::: {{#{feature_target_id} .{sub("^[.]", "", feature_target_class)} .folder-collapse .collapse}}
::::::: {{.folder-body-container}}
[GitHub issue](https://github.com/Gilead-BioStats/gsm.app/issues/{feature$issue})

{acceptance_criteria_qmd}
:::::::
::::::::
:::::::::
'
  )
})

cat(qmd_output, sep = "\n\n")
```

<script>
function handleHash() {
  const hash = window.location.hash;
  if (!hash) return; // Do nothing if there's no hash

  // Construct a valid selector with the "anchor-" prefix
  const selector = '#anchor-' + hash.substring(1);
  const featureDiv = document.querySelector(selector);

  if (featureDiv) {
    const header = featureDiv.querySelector('.folder-header');
    if (header) {
      const isExpanded = header.getAttribute('aria-expanded') === 'true';
      
      // 1. Open the folder only if it's not already expanded
      if (!isExpanded) {
        header.click();
      }

      // 2. Scroll the element into view
      featureDiv.scrollIntoView({ behavior: 'smooth', block: 'start' });
    }
  }
}

// Run the function when the hash changes
window.addEventListener('hashchange', handleHash);

// Run the function on initial page load
document.addEventListener('DOMContentLoaded', handleHash);
</script>
